A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-08-07 survey.tex
casing + bio

Another way  to measure the complexity of the game is  by looking at the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30$ to $300$, and $d \approx 150$ to $200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the average number of actions each unit can execute. 
{\color{blue}
Estimating the value of $u$ is not easy, since the number of actions a unit can execute is highly dependent on the context. Let us make the following assumptions: 1) at most 16 enemy units will be in range of a friendly unit (larger values are possible, but unlikely), 2) when an AI plays StarCraft, it only makes sense to consider movement in the 8 cardinal directions per unit (instead of assuming that the player can issue a ``move'' command to anywhere in the map at any point in time), 3) for ``build'' actions, we consider that SCVs (terran worker units) only build in their current location (otherwise, if they need to move, we consider that as first issuing a ``move'' action, and then a ``build''), and 4) let's consider only the {\em Terran} race. With those assumptions, units in StarCraft can execute between 1 (units like ``Supply Depots'', whose only action is to be ``idle'') to 43 actions (Terran ``Ghosts''), with typical values around 20 to 30. Now, if we have in mind that actions have cool-down times, and thus not all units can execute all of the actions at every frame, we can take a conservative estimation of about $10$ possible actions per unit per game frame. This results in a conservative estimate for the branching factor between $b \in [10^{50},10^{200}]$, only considering units (ignoring the actions buildings can execute). 
}
%Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$.
Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 frames per second).

@@

% POTENTIAL FIELDS:
Potential fields and influence maps have been found to be useful techniques for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multiagent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

Tournament 3 was a tech-limited StarCraft game on a single known map with no fog-of-war enforced. Players
were only allowed to choose the Protoss race, with no late game units allowed. 8 bots
faced off in this double-elimination tournament with mimicbot taking first place over botnik in the final.
As this was a perfect information variant of StarCraft, mimicbot adopted a strategy of
``mimic its opponent's build order, gaining an economic advantage whenever possible'' which
worked quite well.

@@

Results are shown in Table \ref{tab:aiide2010}. The team that won was Overmind\footnote{\url{http://overmind.cs.berkeley.edu}}, from University of California, Berkeley. Using the Zerg race, their
strategy was to defend early aggression with zergling units while amassing mutalisk units,
which they used to contain and eventually defeat their opponents. The mutalisk is a very fast and agile
flying unit which is able to attack while moving with no drawback, which makes them quite a powerful
unit when controlled by a computer. Overmind used a potential-field
based micro-management system to guide their mutalisks, which led them to victory. Krasi0 came in 2nd place
with a standard defensive Terran opening strategy that transitioned into mech play in the late game.

@@

As the CIG competition games were executed manually due to
a lack of available software (the AIIDE program was not yet
available at that time), the organizers separated the ten entries into
two brackets. In each bracket of 5 bots, a round-robin
tournament was held with 10 repetitions per pairing, resulting
in 40 games per bot.
The 5 maps chosen for the first round were selected from the pool
of well-known league play maps found on the internet:
\emph{(2)MatchPoint 1.3}, \emph{(4)Fighting Spirit 1.3}, 
\emph{iCCupdestination 1.1}, \emph{iCCup gaia}, and 
\emph{iCCup great barrier reef}. Each bot pairing played
on every map twice, with switched starting positions.

@@

>>>>

Another way  to measure the complexity of the game is  by looking at the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30$ to $300$, and $d \approx 150$ to $200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the average number of actions each unit can execute. 
{\color{blue}
Estimating the value of $u$ is not easy, since the number of actions a unit can execute is highly dependent on the context. Let us make the following assumptions: 1) at most 16 enemy units will be in range of a friendly unit (larger values are possible, but unlikely), 2) when an AI plays StarCraft, it only makes sense to consider movement in the 8 cardinal directions per unit (instead of assuming that the player can issue a ``move'' command to anywhere in the map at any point in time), 3) for ``build'' actions, we consider that SCVs (Terran worker units) only build in their current location (otherwise, if they need to move, we consider that as first issuing a ``move'' action, and then a ``build''), and 4) let's consider only the {\em Terran} race. With those assumptions, units in StarCraft can execute between 1 (units like ``Supply Depots'', whose only action is to be ``idle'') to 43 actions (Terran ``Ghosts''), with typical values around 20 to 30. Now, if we have in mind that actions have cool-down times, and thus not all units can execute all of the actions at every frame, we can take a conservative estimation of about $10$ possible actions per unit per game frame. This results in a conservative estimate for the branching factor between $b \in [10^{50},10^{200}]$, only considering units (ignoring the actions buildings can execute). 
}
%Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$.
Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 frames per second).

@@

% POTENTIAL FIELDS:
Potential fields and influence maps have been found to be useful techniques for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

Tournament 3 was a tech-limited StarCraft game on a single known map with no fog-of-war enforced. Players
were only allowed to choose the Protoss race, with no late game units allowed. 8 bots
faced off in this double-elimination tournament with MimicBot taking first place over Botnik in the final.
As this was a perfect information variant of StarCraft, MimicBot adopted a strategy of
``mimic its opponent's build order, gaining an economic advantage whenever possible'' which
worked quite well.

@@

Results are shown in Table \ref{tab:aiide2010}. The team that won was Overmind\footnote{\url{http://overmind.cs.berkeley.edu}}, from University of California, Berkeley. Using the Zerg race, their
strategy was to defend early aggression with zergling units while amassing mutalisk units,
which they used to contain and eventually defeat their opponents. The mutalisk is a very fast and agile
flying unit which is able to attack while moving with no drawback, which makes them quite a powerful
unit when controlled by a computer. Overmind used a potential-field
based micro-management system to guide their mutalisks, which led them to victory. Krasi0 came in 2nd place
with a standard defensive Terran opening strategy that transitioned into ``mech'' play in the late game.

@@

As the CIG competition games were executed manually due to
a lack of available software (the AIIDE program was not yet
available at that time), the organizers separated the ten entries into
two brackets. In each bracket of 5 bots, a round-robin
tournament was held with 10 repetitions per pairing, resulting
in 40 games per bot.
The 5 maps chosen for the first round were selected from the pool
of well-known league play maps found on the Internet:
\emph{(2)MatchPoint 1.3}, \emph{(4)Fighting Spirit 1.3}, 
\emph{iCCup Destination 1.1}, \emph{iCCup Gaia}, and 
\emph{iCCup Great Barrier Reef}. Each bot pairing played
on every map twice, with switched starting positions.

@@

% \begin{IEEEbiographynophoto}{Gabriel Synnaeve}
% is a post-doc in the Laboratory of Cognitive Science and Psycholinguistics
% (LSCP) at Ecole Normale Sup\'{e}rieure in Paris, France. Previously, he
% worked on Bayesian modeling in game AI research at INRIA (Grenoble) and
% Coll`{e}ge de France (Paris). He obtained his PhD in 2012 from Grenoble
% University, France.
% \end{IEEEbiographynophoto}
